Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-06-13 thirdPaper/draft/typeConeAssociahedra.tex
Changes to section 3

% math special letters
\newcommand{\R}{\mathbb{R}} % reals
\newcommand{\Q}{\mathbb{Q}} % rationals
\newcommand{\N}{\mathbb{N}} % naturals
\newcommand{\Z}{\mathbb{Z}} % integers
\newcommand{\C}{\mathbb{C}} % complex
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\HH}{\mathbb{H}} % hyperplane
\newcommand{\K}{k} % field
\newcommand{\f}[1]{{\mathfrak{#1}}} % mathfrak letters
\newcommand{\cal}[1]{{\mathcal{#1}}} % call letters
\renewcommand{\b}[1]{{\boldsymbol{#1}}} % bold letters
\newcommand{\h}{\widehat} % hat letters

@@

Finally, we use a similar approach to study the space of polytopal realizations of the $\b{g}$-vector fans of other generalizations of the associahedron.
For non-kissing complexes (a.k.a. support $\tau$-tilting complexes) of gentle algebras, we show that the space of realizations of the non-kissing fan is simplicial when the gentle bound quiver is brick and $2$-acyclic, and we describe in this case its facet-defining inequalities in terms of mesh mutations.
For graphical nested complexes, we show that the space of realizations of the nested fan is simplicial only in the case of the associahedron, and we describe its facet-defining inequalities in general.

@@

The associahedron appears as a fundamental structure in several mathematical theories, such as moduli spaces and topology~\cite{Stasheff, Keller-AinfinityAlgebras}, operads and rewriting theory~\cite{Street, MasudaThomasTonksVallette}, combinatorial Hopf algebras~\cite{LodayRonco, ChatelPilaud, Pilaud-brickAlgebra}, diagonal harmonics~\cite{BergeronPrevilleRatelle, PrevilleRatelleViennot}, mathematical physics~\cite{ArkaniHamedBaiHeYan, BanerjeeLaddhaRaman-StokesPolytopes, Raman-PositiveGeometry}, etc.
Another striking illustration of the ubiquity of the associahedron lies in the theory of cluster algebras introduced by S.~Fomin and A.~Zelevinsky in~\cite{FominZelevinsky-ClusterAlgebrasI} with motivation coming from total positivity and canonical bases.
Cluster algebras are commutative rings generated by variables obtained from an initial seed by the discrete dynamical process of seed mutations.
Finite type cluster algebras, whose mutation graph is finite, were classified in~\cite{FominZelevinsky-ClusterAlgebrasII} using the Cartan-Killing classification for crystallographic root systems.
Finite type cluster complexes were geometrically realized first by F.~Chapoton, S.~Fomin and A.~Zelevinsky~\cite{ChapotonFominZelevinsky} based on the $\b{d}$-vector fans~\cite{FominZelevinsky-YSystems, FominZelevinsky-ClusterAlgebrasII} of the bipartite initial seed, then by~C.~Hohlweg, C.~Lange and H.~Thomas~\cite{HohlwegLangeThomas} based on the Cambrian fans~\cite{Reading-CambrianLattices, ReadingSpeyer} of acyclic initial seeds, and finally by C.~Hohlweg, V.~Pilaud and S.~Stella~\cite{HohlwegPilaudStella} based on the $\b{g}$-vector fans~\cite{FominZelevinsky-ClusterAlgebrasIV} with respect to an arbitrary initial seed.
The resulting polytopes are known as generalized associahedra, and the classical associahedra corresponds to the cluster algebras~of~type~$A$.

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particules, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinetic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\cal{A}$ and consider the real euclidean space~$\R^\cal{V}$ indexed by the set~$\cal{V}$ of cluster variables of~$\cal{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\cal{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\cal{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R^{\cal{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant of~$\R^\cal{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
A by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$ is a simplicial cone.

@@

\medskip
\enlargethispage{.1cm}
Besides revisiting the construction of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and extending it to any initial seed (acyclic or not) in any finite type cluster algebra (simply-laced or not), our reversed approach is also successful when applied to the $\b{g}$-vector fans of other families of generalizations of the associahedron.
In the present paper, we explore specifically two families of such generalizations.

@@

A \defn{polyhedral fan} is a collection~$\Fan$ of polyhedral cones such that
\begin{itemize}
\item if~$C \in \Fan$ and~$F$ is a face of~$C$, then~$F \in \Fan$,
\item the intersection of any two cones of~$\Fan$ is a face of both.
\end{itemize}
A fan is \defn{simplicial} if all its cones are, \defn{complete} if the union of its cones covers the ambient space~$\R^d$, and \defn{essential} if it contains the cone~$\{\b{0}\}$. Note that every complete fan is the product of an essential fan with its linearity space.
For two fans~$\Fan, \cal{G}$ in~$\R^d$, we say that~$\Fan$ \defn{refines}~$\cal{G}$ (and that~$\cal{G}$ \defn{coarsens}~$\Fan$) if every cone of~$\Fan$ is contained in a cone of~$\cal{G}$.

@@

\begin{proof}
Let $\cal{G}$ be a simplicial refinement of $\Fan$. By~\cref{prop:non-simplicial}, the type cone of any coarsening of $\Fan$ is a relatively open cell of the hyperplane arrangement defined by adjacent pairs of~$\cal{G}$. The type cone of the coarsening is not empty whenever it is polytopal. When non-empty, distinct coarsenings give rise to distinct cells (one will have an equality that is a strict inequality in the other), and all lie in the closure of~$\typeCone(\Fan)$ (since they are all in the closure of all the defining half-spaces).
Thus every coarsening describes a relatively open face of~$\ctypeCone(\Fan)$.

@@

\item Mergings induced by extremal adjacent pairs are not the only ones that can take place in the boundary of~$\ctypeCone(\Fan)$. Indeed, if $\cal{G}$ is the coarsening defining a lower-dimensional face~$\typeCone(\cal{G})$ of~$\ctypeCone(\Fan)$, then to get $\Fan$ we have to perform all the coarsenings given by facets of~$\ctypeCone(\Fan)$ containing~$\typeCone(\cal{G})$. But we might have to perform more coarsenings arising from exchangeable pairs  $\{\R_{\ge0}\rays, \R_{\ge0}\rays'\}$ whose inequality is redundant for~$\ctypeCone(\Fan)$ but tight for all points in~$\typeCone(\cal{G})$. This situation arises when, in order to have a polytopal polyhedral fan, more pairs of adjacent cones must be merged.

\item Even if $\Fan$ is essential, not all its polytopal coarsenings need to be. Indeed, it can happen that there is a certain polytopal coarsening $\cal{G}$ in which all the cells have a common linearity space. In this case, $\cal{G}$ is the normal fan of a lower dimensional polytope (whose codimension is the dimension of the linearity space of $\cal{G}$). For this to happen, it is necessary (but not sufficient) that there is a positive linear dependence between some rays of $\Fan$. 
\end{itemize}
\end{remark}

@@

The linear dependences of \cref{lem:linearDependencegvectorsCA} provide a redundant description of the type cone of the cluster fan~$\gvectorFan[\B_\circ]$.
We denote by~$\b{n}(\B_\circ, x, x')$ the normal vector of the inequality of the type cone corresponding to two exchangeable cluster variables~$x$ to~$x'$ (it is well defined by \cref{prop:uniqueExchangePropertyCA}).
In other words, depending on which of the two linear dependences of \cref{lem:linearDependencegvectorsCA} holds, we have
\[
\b{n}(\B_\circ, x, x') \eqdef \b{f}_x + \b{f}_{x'} - \sum_{\substack{y \in \cluster \cap \cluster' \\ b_{xy} < 0}} -b_{xy} \, \b{f}_y
\qquad\text{or}\qquad
\b{n}(\B_\circ, x, x') \eqdef \b{f}_x + \b{f}_{x'} - \sum_{\substack{y \in \cluster \cap \cluster' \\ b_{xy} > 0}} b_{xy} \, \b{f}_y,
\]
where~$(\b{f}_x)_{x \in \principalVariables}$ denotes the canonical basis of~$\R^{\principalVariables}$.
We obtain the following statement.

@@

The following statement is proved in \cref{coro::number of positive meshes}.

\begin{lemma}
\label{lem:bijectionMeshMutations}
We have that~$|\principalVariables| = |\meshes| + n$.
\end{lemma}

The following statement describes the linear dependence in the mesh mutations. It follows from \cref{lem:linearDependencegvectorsCA,def:meshMutation}.

@@

\section{Relations for~$\b{g}$-vectors in finite type cluster algebras \\ via~$2$-Calabi--Yau triangulated categories}
\label{sec:clusterCategories}

In~\cite{Auslander1984}, M.~Auslander showed that the relations in the Grothendieck group of an Artin algebra are generated by the ones given by almost-split sequences precisely when the algebra is of finite representation type. Inspired by this result, we will show that a similar statement holds (see \cref{theo::relations-g-vecteurs}) for the~$\b{g}$-vectors (or indices) of objects in a~$2$-Calabi--Yau triangulated category.

@@

Now,~$(\susp T)\cat(\susp^{-1} X, \susp X)$ is a sub-module of~$\cat(\susp^{-1} X, \susp X)$. Therefore, if~$(\susp T)\cat(\susp^{-1} X, \susp X)$ is non-zero, then it contains the one-dimensional socle of~$\cat(\susp^{-1} X, \susp X)$, and thus contains~$h$.  
By~\cite{Palu},
\[
(\susp T)\cat(\susp^{-1} X, \susp X) \cong D\cat(X, X)/(\susp T). 
\]
The identity morphism of~$X$ is not in~$(\susp T)$, since~$X$ is not in~$\add(\susp T)$. Thus the right-hand side is non-zero, and so neither is the left-hand side. By the above, this implies~${h \in (\susp T)\cat(\susp^{-1} X, \susp X)}$, which finishes the proof.
\end{proof}

\begin{lemma}
\label{lemm::bilinear-form}
Let~$A$ and~$B$ be two indecomposable objects of~$\cat$. \pg{introduce general notation for redisue field}
\begin{enumerate}
\item If~$A \notin \add(\susp T)$, then
\[
\langle \ell_A, B \rangle = 
\begin{cases}
0 & \textrm{if~$A\ncong B$;} \\
\dim_{\field}\field_A & \textrm{if~$A\cong B$.}
\end{cases}
\]

@@

\begin{proof}
We only prove the first assertion; the second one is proved dually. Assume that ${A\notin\add(\susp T)}$. Let
\[
A \xrightarrow{f} E \xrightarrow{} \susp^{-1} A \xrightarrow{h} \susp A
\]
be an almost-split triangle. By \cref{lemm::almost-split}, the morphism~$h$ is in~$(\susp T)$. Applying the functor ${F=\cat(T,-)}$, we get an exact sequence
\[
FA\xrightarrow{Ff} FE \xrightarrow{} F\susp^{-1} A \xrightarrow{} 0.
\]
Applying now the functors~$\cat(-, B)$ and~$\Hom{\Lambda}(-, FB)$, we get a commutative diagram whose rows are exact sequences and whose vertical maps are surjective by \cref{prop::functor-F}.
\[
\xymatrix{ & \cat(\susp^{-1} A, B)\ar[r]\ar@{->>}[d] & \cat(E, B)\ar[r]^{f^*}\ar@{->>}[d] & \cat(A,B)\ar[r]\ar@{->>}[d] & \coker(f^*)\ar[r]\ar@{->>}[d] & 0 \\
0 \ar[r] & \Hom{\Lambda}(F\susp^{-1} A, FB)\ar[r] & \Hom{\Lambda}(FE, FB)\ar[r]^{Ff^*} & \Hom{\Lambda}(FA, FB)\ar[r] & \coker(Ff^*)\ar[r] & 0.
}
\]

If~$B\ncong A$, then the definition of an almost-split triangle implies that~$f^*$ is surjective. Thus $\coker(f^*)=0$, so that~$\coker(Ff^*) = 0$, and by additivity of the dimension in exact sequences, we get that~$\langle \ell_A, B \rangle =0$.

@@

\begin{proposition}
\label{prop::free-set}
\begin{enumerate}
\item The set~$\set{[\ell_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i \in [n]}$ is free in~$\Ksp(\cat)$.
\item The set~$\set{[r_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i \in [n]}$ is free in~$\Ksp(\cat)$.
\end{enumerate}
\end{proposition}

\begin{proof}
We only prove (1); the proof of (2) is similar. Assume that~
\[
x= \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \lambda_A \ell_A + \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
\]
Then~$\langle x, [A] \rangle = 0$ for all~$A\in\ind(\cat) \ssm \add(\susp T)$. But~$\langle x, [A] \rangle = \lambda_A$ by \cref{lemm::bilinear-form}. Thus~
\[
x = \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
\]
But the~$[\susp T_i]$ are linearly independent in~$\Ksp(\cat)$. Thus~$\lambda_i = 0$ for all~$i \in [n]$. This finishes the proof.
\end{proof}

@@

\begin{corollary}
If~$\ind(\cat)$ is finite, then the set
\[
\set{[\ell_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i \in [n]}
\]
is a basis of~$\Ksp(\cat)$.
\end{corollary}

\begin{proof}
By \cref{prop::free-set}, the set is free. It suffices to prove that it generates~$\Ksp(\cat)$. Let~${x \in \Ksp(\cat)}$. Consider
\[
z = x - \sum_{A \in \ind{\cat}} \frac{\langle x, A \rangle}{\langle \ell_A, A\rangle} \ell_A.
\]
Then for any~$B \in \ind(\cat)$, we have that~$\langle z, [B] \rangle = 0$. By \cref{lemm::test-susp-T}, this implies that ${z \in \Ksp \big( \add(\susp T) \big)}$, and finishes the proof.
\end{proof}

@@

Now,~$[T]+[\susp T]$ is in the kernel of~$\b{g}$, but~$\langle [T] + [\susp T], [B] \rangle = \langle [T], [B] \rangle = 0$ if and only if ${B \in \add(\susp T)}$. Thus~$\ind(\cat)$ has to be finite.
\end{proof}

@@

\begin{definition}
\label{defn::positive mutation}
Let~$x\in\cluster$ be a cluster variable in a given cluster and let~$x' \in \cluster'$ be obtained by mutating~$\cluster$ at~$x$.
The pair~$(x,x')$ is said to be a \defn{positive mutation} with respect to some initial cluster~$\cluster_\circ$ if there are~$X,Y\in \ind(\cat)$ such that~$\CC(X)=x$,~$\CC(Y)=x'$,~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$ and, for any non-split triangle
\[
X\xrightarrow{} E \xrightarrow{} Y \xrightarrow{h} \susp X 
\]
(this triangle is unique up to isomorphism) we have~$h\in (\susp T)$, where~$T\in\cat$ is any basic cluster-tilting object such that~$\CC(T)=\cluster_\circ$.

@@

\begin{lemma}
\label{lemm::exaclty one mesh mutation}
For every non-initial cluster variable, there exists precisely one pair~$(x,x')$ such that there exists a positive mesh mutation changing~$x$ into~$x'$.
\end{lemma}

\begin{proof}
Up to isomorphism, there is precisely one almost-split triangle of~$\cat$ starting in~$X$, where~$X$ is the object such that~$\phi(X) = x$.  Thus there is precisely one mesh relation By \cref{lemm::almost-split}, the pair~$(x,x')$ is a positive mutation if and only~$X\notin \add(\susp T)$, that is, if and only if~$x$ is not an initial variable.
\end{proof}

\begin{lemma}
\label{lemm::mesh mutations and meshes}
The set of positive mesh mutations~$(x,x')$ of \cref{defn::positive mutation} is in bijection with the set~$\meshes$ of~\cref{defn::positive mutation}.
\end{lemma}

>>>>

% math special letters
\newcommand{\R}{\mathbb{R}} % reals
\newcommand{\Q}{\mathbb{Q}} % rationals
\newcommand{\N}{\mathbb{N}} % naturals
\newcommand{\Z}{\mathbb{Z}} % integers
\newcommand{\C}{\mathbb{C}} % complex
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\HH}{\mathbb{H}} % hyperplane
\newcommand{\K}{k} % field
\newcommand{\f}[1]{{\mathfrak{#1}}} % mathfrak letters
\renewcommand{\c}[1]{{\mathcal{#1}}} % call letters
\renewcommand{\b}[1]{{\boldsymbol{#1}}} % bold letters
\newcommand{\h}{\widehat} % hat letters

@@

Finally, we use a similar approach to study the space of polytopal realizations of the $\b{g}$-vector fans of other generalizations of the associahedron.
For non-kissing complexes (a.k.a. support~$\tau$-tilting complexes) of gentle algebras, we show that the space of realizations of the non-kissing fan is simplicial when the gentle bound quiver is brick and $2$-acyclic, and we describe in this case its facet-defining inequalities in terms of mesh mutations.
For graphical nested complexes, we show that the space of realizations of the nested fan is simplicial only in the case of the associahedron, and we describe its facet-defining inequalities in general.

@@

The associahedron appears as a fundamental structure in several mathematical theories, such as moduli spaces and topology~\cite{Stasheff, Keller-AinfinityAlgebras}, operads and rewriting theory~\cite{}, combinatorial Hopf algebras~\cite{LodayRonco, ChatelPilaud, Pilaud-brickAlgebra}, diagonal harmonics~\cite{}, mathematical physics~\cite{ArkaniHamedBaiHeYan, BanerjeeLaddhaRaman-StokesPolytopes, Raman-PositiveGeometry}, etc.
Another striking illustration of the ubiquity of the associahedron lies in the theory of cluster algebras introduced by S.~Fomin and A.~Zelevinsky in~\cite{FominZelevinsky-ClusterAlgebrasI} with motivation coming from total positivity and canonical bases.
Cluster algebras are commutative rings generated by variables obtained from an initial seed by the discrete dynamical process of seed mutations.
Finite type cluster algebras, whose mutation graph is finite, were classified in~\cite{FominZelevinsky-ClusterAlgebrasII} using the Cartan-Killing classification for crystallographic root systems.
Finite type cluster complexes were geometrically realized first by F.~Chapoton, S.~Fomin and A.~Zelevinsky~\cite{ChapotonFominZelevinsky} based on the $\b{d}$-vector fans~\cite{FominZelevinsky-YSystems, FominZelevinsky-ClusterAlgebrasII} of the bipartite initial seed, then by~C.~Hohlweg, C.~Lange and H.~Thomas~\cite{HohlwegLangeThomas} based on the Cambrian fans~\cite{Reading-CambrianLattices, ReadingSpeyer} of acyclic initial seeds, and finally by C.~Hohlweg, V.~Pilaud and S.~Stella~\cite{HohlwegPilaudStella} based on the $\b{g}$-vector fans~\cite{FominZelevinsky-ClusterAlgebrasIV} with respect to an arbitrary initial seed.
The resulting polytopes are known as generalized associahedra, and the classical associahedra corresponds to the cluster algebras of type~$A$.

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particules, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinetic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\c{A}$ and consider the real euclidean space~$\R^\c{V}$ indexed by the set~$\c{V}$ of cluster variables of~$\c{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\c{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\c{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R^{\c{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant of~$\R^\c{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\c{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
A by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\c{A}$ with respect to the initial acyclic seed~$\seed_\circ$ is a simplicial cone.

@@

\medskip
Besides revisiting the construction of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and extending it to any initial seed (acyclic or not) in any finite type cluster algebra (simply-laced or not), our reversed approach is also successful when applied to the $\b{g}$-vector fans of other families of generalizations of the associahedron.
In the present paper, we explore specifically two families of such generalizations.

@@

A \defn{polyhedral fan} is a collection~$\Fan$ of polyhedral cones such that
\begin{itemize}
\item if~$C \in \Fan$ and~$F$ is a face of~$C$, then~$F \in \Fan$,
\item the intersection of any two cones of~$\Fan$ is a face of both.
\end{itemize}
A fan is \defn{simplicial} if all its cones are, \defn{complete} if the union of its cones covers the ambient space~$\R^d$, and \defn{essential} if it contains the cone~$\{\b{0}\}$. Note that every complete fan is the product of an essential fan with its linearity space.
For two fans~$\Fan, \c{G}$ in~$\R^d$, we say that~$\Fan$ \defn{refines}~$\c{G}$ (and that~$\c{G}$ \defn{coarsens}~$\Fan$) if every cone of~$\Fan$ is contained in a cone of~$\c{G}$.

@@

\begin{proof}
Let $\c{G}$ be a simplicial refinement of $\Fan$. By~\cref{prop:non-simplicial}, the type cone of any coarsening of $\Fan$ is a relatively open cell of the hyperplane arrangement defined by adjacent pairs of~$\c{G}$. The type cone of the coarsening is not empty whenever it is polytopal. When non-empty, distinct coarsenings give rise to distinct cells (one will have an equality that is a strict inequality in the other), and all lie in the closure of~$\typeCone(\Fan)$ (since they are all in the closure of all the defining half-spaces).
Thus every coarsening describes a relatively open face of~$\ctypeCone(\Fan)$.

@@

\item Mergings induced by extremal adjacent pairs are not the only ones that can take place in the boundary of~$\ctypeCone(\Fan)$. Indeed, if $\c{G}$ is the coarsening defining a lower-dimensional face~$\typeCone(\c{G})$ of~$\ctypeCone(\Fan)$, then to get $\Fan$ we have to perform all the coarsenings given by facets of~$\ctypeCone(\Fan)$ containing~$\typeCone(\c{G})$. But we might have to perform more coarsenings arising from exchangeable pairs  $\{\R_{\ge0}\rays, \R_{\ge0}\rays'\}$ whose inequality is redundant for~$\ctypeCone(\Fan)$ but tight for all points in~$\typeCone(\c{G})$. This situation arises when, in order to have a polytopal polyhedral fan, more pairs of adjacent cones must be merged.

\item Even if $\Fan$ is essential, not all its polytopal coarsenings need to be. Indeed, it can happen that there is a certain polytopal coarsening $\c{G}$ in which all the cells have a common linearity space. In this case, $\c{G}$ is the normal fan of a lower dimensional polytope (whose codimension is the dimension of the linearity space of $\c{G}$). For this to happen, it is necessary (but not sufficient) that there is a positive linear dependence between some rays of $\Fan$. 
\end{itemize}
\end{remark}

@@

The linear dependences of \cref{lem:linearDependencegvectorsCA} provide a redundant description of the type cone of the cluster fan~$\gvectorFan[\B_\circ]$.
We denote by~$\b{n}(\B_\circ, x, x')$ the normal vector of the inequality of the type cone corresponding to two exchangeable cluster variable~$x$ to~$x'$ (it is well defined by \cref{prop:uniqueExchangePropertyCA}).
In other words, depending on which of the two linear dependences of \cref{lem:linearDependencegvectorsCA} holds, we have
\[
\b{n}(\B_\circ, x, x') \eqdef \b{f}_x + \b{f}_{x'} - \sum_{\substack{y \in \cluster \cap \cluster' \\ b_{xy} < 0}} -b_{xy} \, \b{f}_y
\qquad\text{or}\qquad
\b{n}(\B_\circ, x, x') \eqdef \b{f}_x + \b{f}_{x'} - \sum_{\substack{y \in \cluster \cap \cluster' \\ b_{xy} > 0}} b_{xy} \, \b{f}_y,
\]
where~$(\b{f}_x)_{x \in \principalVariables}$ denotes the canonical basis of~$\R^{\principalVariables}$.
We obtain the following statement.

@@

%\vincent{Link to \cref{sec:clusterCategories}.}

\begin{lemma}
\label{lem:bijectionMeshMutations}
We have that~$|\principalVariables| = |\meshes| + n$.
\end{lemma}
\begin{proof}
 We postpone the proof until Corollary~\ref{coro::number of positive meshes}.
\end{proof}

The following statement describes the linear dependence in the mesh mutations.

@@

\section{Relations for~$\b{g}$-vectors in finite type cluster algebras via~$2$-Calabi--Yau triangulated categories}
\label{sec:clusterCategories}

In~\cite{Auslander1984}, M.~Auslander showed that the relations in the Grothendieck group of an Artin algebra are generated by the ones given by almost-split sequences precisely when the algebra is of finite representation type. Inspired by this result, we will show that a similar statement holds (\cref{theo::relations-g-vecteurs}) for the~$\b{g}$-vectors (or indices) of objects in a~$2$-Calabi--Yau triangulated category.

@@

Now,~$(\susp T)\cat(\susp^{-1} X, \susp X)$ is a sub-module of~$\cat(\susp^{-1} X, \susp X)$. Therefore, if~$(\susp T)\cat(\susp^{-1} X, \susp X)$ is non-zero, then it contains the one-dimensional socle of~$\cat(\susp^{-1} X, \susp X)$, and thus contains~$h$.  
By~\cite{Palu},
\[
(\susp T)\cat(\susp^{-1} X, \susp X) \cong D\cat(X, X)/(\susp T). 
\]
The identity morphism of~$X$ is not in~$(\susp T)$, since~$X$ is not in~$\add(\susp T)$. Thus the right-hand side is non-zero, and so neither is the left-hand side. By the above, this implies that~$h \in (\susp T)\cat(\susp^{-1} X, \susp X)$, which finishes the proof.
\end{proof}

\begin{lemma}
\label{lemm::bilinear-form}
Let~$A$ and~$B$ be two indecomposable objects of~$\cat$. %\pg{introduce general notation for redisue field}
\begin{enumerate}
\item If~$A \notin \add(\susp T)$, then
\[
\langle \ell_A, B \rangle = 
\begin{cases}
0 & \textrm{if~$A\ncong B$;} \\
\dim_{\field}\field_A & \textrm{if~$A\cong B$.}
\end{cases}
\]

@@

\begin{proof}
We only prove the first assertion; the second one is proved dually. Assume that~$A\notin\add(\susp T)$. Let
\[
A \xrightarrow{f} E \xrightarrow{} \susp^{-1} A \xrightarrow{h} \susp A
\]
be an almost-split triangle. By \cref{lemm::almost-split}, the morphism~$h$ is in~$(\susp T)$. Applying the functor~$F=\cat(T,-)$, we get an exact sequence
\[
FA\xrightarrow{Ff} FE \xrightarrow{} F\susp^{-1} A \xrightarrow{} 0.
\]
Applying now the functors~$\cat(-, B)$ and~$\Hom{\Lambda}(-, FB)$, we get a commutative diagram whose rows are exact sequences and whose vertical maps are surjective by \cref{prop::functor-F}.
\[
\xymatrix{ & \cat(\susp^{-1} A, B)\ar[r]\ar@{->>}[d] & \cat(E, B)\ar[r]^{f^*}\ar@{->>}[d] & \cat(A,B)\ar[r]\ar@{->>}[d] & \coker(f^*)\ar[r]\ar@{->>}[d] & 0 \\
0 \ar[r] & \Hom{\Lambda}(F\susp^{-1} A, FB)\ar[r] & \Hom{\Lambda}(FE, FB)\ar[r]^{Ff^*} & \Hom{\Lambda}(FA, FB)\ar[r] & \coker(Ff^*)\ar[r] & 0.
}
\]

If~$B\ncong A$, then the definition of an almost-split triangle implies that~$f^*$ is surjective. Thus~$\coker(f^*)=0$, so that~$\coker(Ff^*) = 0$, and by additivity of the dimension in exact sequences, we get that~$\langle \ell_A, B \rangle =0$.

@@

\begin{proposition}
\label{prop::free-set}
\begin{enumerate}
\item The set~$\set{[\ell_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i=1, \dots, n}$ is free in~$\Ksp(\cat)$.
\item The set~$\set{[r_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i=1, \dots, n}$ is free in~$\Ksp(\cat)$.
\end{enumerate}
\end{proposition}

\begin{proof}
We only prove (1); the proof of (2) is similar. Assume that~
\[
x= \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \lambda_A \ell_A + \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
\]
Then~$\langle x, [A] \rangle = 0$ for all~$A\in\ind(\cat) \ssm \add(\susp T)$. But~$\langle x, [A] \rangle = \lambda_A$ by \cref{lemm::bilinear-form}. Thus~
\[
x = \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
\]
But the~$[\susp T_i]$ are linearly independent in~$\Ksp(\cat)$. Thus~$\lambda_i = 0$ for all~$i \in \{1, \ldots, n\}$. This finishes the proof.
\end{proof}

@@

\begin{corollary}
If~$\ind(\cat)$ is finite, then the set
\[
\set{[\ell_A]}{A \in \ind(\cat) \ssm \add(\susp T)} \cup \set{[\susp T_i]}{i=1, \ldots, n}
\]
is a basis of~$\Ksp(\cat)$.
\end{corollary}

\begin{proof}
By \cref{prop::free-set}, the set is free. It suffices to prove that it generates~$\Ksp(\cat)$. Let~${x \in \Ksp(\cat)}$. Consider
\[
z = x - \sum_{A \in \ind{\cat}} \frac{\langle x, A \rangle}{\langle \ell_A, A\rangle} \ell_A.
\]
Then for any~$B \in \ind(\cat)$, we have that~$\langle z, [B] \rangle = 0$. By \cref{lemm::test-susp-T}, this implies that~${z \in \Ksp \big( \add(\susp T) \big)}$, and finishes the proof.
\end{proof}

@@

Now,~$[T]+[\susp T]$ is in the kernel of~$\b{g}$, but~$\langle [T] + [\susp T], [B] \rangle = \langle [T], [B] \rangle = 0$ if and only if~$B \in \add(\susp T)$. Thus~$\ind(\cat)$ has to be finite.
\end{proof}

@@

\begin{definition}\label{defn::positive mutation}
Let~$x\in\cluster$ be a cluster variable in a given cluster and let~$x' \in \cluster'$ be obtained by mutating~$\cluster$ at~$x$.
The pair~$(x,x')$ is said to be a \defn{positive mutation} with respect to some initial cluster~$\cluster_\circ$ if there are~$X,Y\in \ind(\cat)$ such that~$\CC(X)=x$,~$\CC(Y)=x'$,~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$ and, for any non-split triangle
 \[
  X\xrightarrow{} E \xrightarrow{} Y \xrightarrow{h} \susp X 
 \]
(this triangle is unique up to isomorphism) we have~$h\in (\susp T)$, where~$T\in\cat$ is any basic cluster-tilting object such that~$\CC(T)=\cluster_\circ$.

@@

\begin{lemma}\label{lemm::exaclty one mesh mutation}
 For every non-initial cluster variable~$x$, there exists precisely one pair~$(x,x')$ such that there exists a positive mesh mutation changing~$x$ into~$x'$.
\end{lemma}
\begin{proof}
 Up to isomorphism, there is precisely one almost-split triangle of~$\cat$ starting in~$X$, where~$X$ is the object such that~$\phi(X) = x$.  Thus there is precisely one mesh mutation. By Lemma~\ref{lemm::almost-split}, the pair~$(x,x')$ is a positive mutation if and only~$X\notin \add(\susp T)$, that is, if and only if~$x$ is not an initial variable.
\end{proof}
